Augmentation of a Model Historical Harpsichord Mechanism for Haptic Enabled Interaction

### 2025-01-18 src/article.tex
double label
fig to ref

The haptic key project by Timmermans et al.~\cite{Timmermans2020} began
with the model of a piano key. A 3-key cross-sectional model of a
harpischord mechanism \figure{3key} was provided to the NEMUS project by
San Colombano as a test bed for sensors and as a proof of concept for a
larger scale interface. Each key of model has 2 sets of strings that are
under tension but not at any specific musical pitch. The strings are
there to provide same resistance to the jack plectra as would be found
on an instrument. The mechanism is in the italian style of the XVI-XIX
century with 2 jacks per key with corresponding strings tuned in unison
\footnote{"8'-8' È la dispozione normale dei clavicembali
italiani" \cite{Tagliavini1987}}. A suitable sensing system was found
for reliably detecting when jack had plucked a string independently and
sending out a corresponding MIDI message to a connected sound synthesis
engine. A full scale 49-key scale model was commissioned with and then
built by luthier Roberto Livi in the same style as the 1547 Trasuntino
\figure{teaser}. The sensor system was installed into the full scale
model and then setup for exhibition at San Colombano. Visitors are
invited to play the interface which currently triggers samples on a
connected sound engine. The paper looks at the design decisions,
processes, and considerations to be made when undertaking a project of
this kind.

\begin{figure}
  \label{}
  \centering
  \includegraphics[width=\linewidth]{images/side_w-o_sensors.png} 
  \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023} \Description{A 3-Key Model of a Harpsichord Mechanism created Graziano Bandini in Bologna. A sidevew of the model which is a cross-section of what is a common
mechanism for an italian style harpischord of the XVI century.} 
\label{fig:3key}
\end{figure}

>>>>

The haptic key project by Timmermans et al.~\cite{Timmermans2020} began
with the model of a piano key. A 3-key cross-sectional model of a
harpischord mechanism \ref{3key} was provided to the NEMUS project by
San Colombano as a test bed for sensors and as a proof of concept for a
larger scale interface. Each key of model has 2 sets of strings that are
under tension but not at any specific musical pitch. The strings are
there to provide same resistance to the jack plectra as would be found
on an instrument. The mechanism is in the italian style of the XVI-XIX
century with 2 jacks per key with corresponding strings tuned in unison
\footnote{"8'-8' È la dispozione normale dei clavicembali
italiani" \cite{Tagliavini1987}}. A suitable sensing system was found
for reliably detecting when jack had plucked a string independently and
sending out a corresponding MIDI message to a connected sound synthesis
engine. A full scale 49-key scale model was commissioned with and then
built by luthier Roberto Livi in the same style as the 1547 Trasuntino
\ref{teaser}. The sensor system was installed into the full scale
model and then setup for exhibition at San Colombano. Visitors are
invited to play the interface which currently triggers samples on a
connected sound engine. The paper looks at the design decisions,
processes, and considerations to be made when undertaking a project of
this kind.

\begin{figure}  
  \centering
  \includegraphics[width=\linewidth]{images/side_w-o_sensors.png} 
  \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023} \Description{A 3-Key Model of a Harpsichord Mechanism created Graziano Bandini in Bologna. A sidevew of the model which is a cross-section of what is a common
mechanism for an italian style harpischord of the XVI century.} 
\label{fig:3key}
\end{figure}
